---
title: "Check the completeness of follow-up in the ADRD population"
output: github_document
---

The dataset `ADRDcohort.fst` was directly generated from the denominator files, which follow the ADRD people till their death or the end of 2016.

## Setup and load data

```{r echo=FALSE}
rm(list = ls())
gc()

library(data.table)
setDTthreads(threads = 0)
library(dplyr)
library(NSAPHutils)
library(fst)

# setwd("/nfs/home/S/shd968/shared_space/ci3_shd968/dementia")
dir_ADRDcohort <-  "/nfs/home/S/shd968/shared_space/ci3_shd968/dementia/data/"
dir_enrolledInfo <-  "/nfs/home/S/shd968/shared_space/ci3_shd968/dementia/data/"

ADRDcohort <- read_fst(paste0(dir_ADRDcohort, "ADRDcohort.fst"), as.data.table = T)
enrolledInfo <- fread(paste0(dir_enrolledInfo, "EnrolledInfo.csv"))
```

## Check the completeness of follow-up

Our final dataset should:

- be one row per person-year (long format)
- have follow-up info for every year starting from `firstADRDyr` till the end of study period (2016) or death or leaving the US, whichever comes first (though we do not need the info for `year==firstADRDyr` in survival analysis, we do need it for correcting index event bias)

### 1. basic info of original ADRD dataset
```{r}
# number of subjects in raw cohort
uniqueN(ADRDcohort[,qid])
# number of subjects in cohorts after removing NAs
uniqueN(na.omit(ADRDcohort)[,qid])
```

Those who admitted for the first time with ADRD code and died in the same year don't contribute to the risk set. Delete those

```{r}
A <- ADRDcohort[, .(end_yr = max(year)), by = qid]
A <- A[enrolledInfo, on = .(qid=QID)]
no_contribution <- subset(A, end_yr==firstADRDyr)
```

```{r}
temp <- na.omit(ADRDcohort[!(qid %in% no_contribution[,qid])]) # remove all the NAs and those without contribution to risk set
sum_temp <- temp[, .(start_yr = min(year),
                     end_yr = max(year),
                     count = uniqueN(year)), by = qid]
sum_temp <- merge(sum_temp, enrolledInfo, by.x = "qid", by.y = "QID", all.x = TRUE)
```

We constructed a temporary dataset named `temp` which is a subset of `ADRDcohort` after removing NA (not remove rows with `year==firstADRDyr`), and summarize each person as one row in total `r dim(sum_temp)[1]` subjects:

- generated `start_yr` as the minimum of calendar year (should equal to `firstADRDyr`)
- `end_yr` as the maximum of calendar year
- `count` as the count number of unique calendar year for each subject. 

We also merged the enroll information (`firstADRDyear`) to `sum_temp`. `firstADRDyr` indicates the year that subjects should start to be followed-up.

`sum_temp` is a one-row-per-person dataset, looks like:

```{r echo=FALSE}
head(sum_temp)
```

```{r echo=FALSE}
cat("the number of subjects in ADRD cohort (removed NAs, and non-contribution subjects) is", dim(sum_temp)[1], "\n")
cat("the number of person-years in ADRD cohort (removed NAs, and non-contribution subjects) is", dim(temp)[1], "\n")
cat("is there any duplication of the combination of `qid` and calendar year: ", any(duplicated(temp[,.(qid,year)])), "\n")
cat("is there any duplication of the combination of `qid` and age: ", any(duplicated(temp[,.(qid,age)])), "\n")
```

### 2. check whether all people were followed-up from the year following firstADRDyr
```{r}
sum_temp[start_yr != (firstADRDyr)]
```

The above subject (`r dim(sum_temp[start_yr != (firstADRDyr+1)])[1]` subjects in total), were not followed-up from the year following firstADRDyr. 

Their info in denominator files:
```{r echo=FALSE}
temp[qid %in% sum_temp[start_yr != firstADRDyr, qid], ]
```

```{r echo=FALSE}
cat("the number of person-years of related subjects is", dim(temp[qid %in% sum_temp[start_yr != firstADRDyr, qid], ])[1], "\n")
```


<!-- ```{r summarize} # dead event summary(temp[qid %in% sum_temp[start_yr != -->
<!-- firstADRDyr, qid],.SD[.N],by=qid][,dead]) # almost the same as the general -->
<!-- population ``` -->
<!-- <!-- #### fix --> -->

<!-- ```{r operature} -->
<!-- # create row for each missing year -->
<!-- op1 <- sum_temp[start_yr != (firstADRDyr+1),.(qid, firstADRDyr, year=start_yr-1)] -->
<!-- op2 <- op1[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op3 <- op2[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op4 <- op3[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op5 <- op4[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op6 <- op5[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op7 <- op6[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op8 <- op7[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op9 <- op8[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op10 <- op9[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op11 <- op10[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op12 <- op11[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op13 <- op12[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op14 <- op13[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op15 <- op14[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- op16 <- op15[year!=firstADRDyr+1,.(qid, firstADRDyr, year=year-1)] -->
<!-- insert <- rbind(op1,op2,op3,op4,op5,op6,op7,op8,op9,op10,op11,op12,op13,op14,op15,op16) -->
<!-- setorder(insert,qid, year) -->
<!-- insert # dataset to be inserted -->
<!-- ``` -->

<!-- ```{r covariates} -->
<!-- # covariates info by zip and year -->
<!-- covariates <- unique(temp[,.(zip, year, mean_bmi, smoke_rate, hispanic, pct_blk, medhouseholdincome, medianhousevalue, poverty, education, popdensity, pct_owner_occ, summer_tmmx, winter_tmmx, summer_rmax, winter_rmax, pm25, no2, ozone, ozone_summer)]) -->
<!-- ``` -->

<!-- ```{r demographic} -->
<!-- setorder(temp, qid, year) -->
<!-- demographic <- temp[,.SD[1],by=qid][,diff_ageyear:=year-age][,.(qid, sex, race, dual, statecode,diff_ageyear)] -->
<!-- demographic -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # get zip info to "insert"" datset -->
<!-- temp[qid %in% insert[,qid],.(qid,zip)][,uniqueN(zip),by=qid][V1!=1,] # check if zip change during follow-up -->
<!-- zipinfo <- temp[qid %in% insert[,qid],.(qid,zip)][,.SD[1],by=qid] # take one of zip -->
<!-- insert <- merge(insert, zipinfo, by="qid", all.x = T) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- insert <- merge(insert, covariates, by=c("year", "zip"), all.x = T) -->
<!-- anyNA(insert) -->

<!-- insert <- merge(insert, demographic, by="qid", all.x=T) -->
<!-- insert[,age:=year-diff_ageyear] -->
<!-- insert[,diff_ageyear:=NULL] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- insert[,`:=`(year_prev=year-1,dead=F)] -->
<!-- ADRDcohort_added <- rbind(insert,temp) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- temp_added <- na.omit(ADRDcohort_added) # remove all the NAs -->
<!-- sum_temp_added <- temp_added[, .(start_yr = min(year), -->
<!--                      end_yr = max(year), -->
<!--                      count = uniqueN(year)), by = qid] -->
<!-- sum_temp_added <- merge(sum_temp_added, enrolledInfo, by.x = "qid", by.y = "QID", all.x = TRUE) -->
<!-- ``` -->

### 3. check whether all people have each year's info during follow-up
```{r}
sum_temp[(end_yr-firstADRDyr+1) != count,]
setorder(temp,qid,year)
# summary(temp[qid%in%sum_temp[(end_yr-firstADRDyr+1) != count,qid],][,.SD[.N],by=qid][,dead])
```

```{r}
## save qid without complete follow-up or no contribution to risk set
omitInfo <- rbind(sum_temp[(end_yr-firstADRDyr+1) != count, .(qid,end_yr, firstADRDyr)], no_contribution)
fwrite(omitInfo, "/nfs/home/S/shd968/shared_space/ci3_shd968/dementia/data/omitInfo.csv")
```

The above subject (`r dim(omitInfo)[1]` subjects in total) do not have each year's info during follow-up. **Consider deleting them**

Their info in denominator files:
```{r echo=FALSE}
temp[qid %in% sum_temp[(end_yr-firstADRDyr+1) != count,][,qid],]
```

### 4. check whether alive people were followed-up till the end of study period (2016) after excluding those without complete follow-up
```{r}
sum_temp[!(qid %in% temp[(dead),qid]), ][!(qid %in% omitInfo[,qid]),end_yr] %>% table()
```

We could see loads of alive subjects weren't followed-up till 2016. This should be considered as right-censored subjects in the analyses.