---
title: "checkFollowupCompleteness"
output:
  md_document:
    variant: markdown_github
---

# Check the completeness of follow-up in the ADRD population

The dataset `ADRDpeople.csv` was directly generated from the denominator files.

## Setup and read in data

```{r echo=FALSE}
rm(list = ls())
gc()

library(data.table)
library(dplyr)

setDTthreads(threads = 0)
setwd("/nfs/home/S/shd968/shared_space/ci3_shd968/dementia")
dir_ADRDpeople <-  "/nfs/home/S/shd968/shared_space/ci3_shd968/dementia/"
dir_enrolledInfo <-  "/nfs/home/S/shd968/shared_space/ci3_shd968/dementia/"

ADRDpeople <- fread(paste0(dir_ADRDpeople, "ADRDpeople.csv"))
enrolledInfo <- fread(paste0(dir_enrolledInfo, "EnrolledInfo.csv"))
```

## Check the completeness of follow-up

Our final dataset should:

- be one row per person-year
- have follow-up info for every year after the year they were hospitalized with ADRD for the first time (`firstADRDyr`) till the end of study period (2016) or death, whichever comes first.

### 1. basic info of original ADRD dataset
```{r}
temp <- na.omit(ADRDpeople)
temp <- temp[year!=firstADRDyr, ][, .(start_yr = min(year),
                                           end_yr = max(year),
                                           count = uniqueN(year)), by = qid]
temp <- merge(temp, enrolledInfo, by.x = "qid", by.y = "QID")
```

We constructed a temporary dataset named `temp` which is a subset of `ADRDpeople` after removing NA and removing rows whose `year==firstADRDyr`.

We generated `start_yr` as the minimum of calendar year and `end_yr` as the maximum of calendar year, also, `count` as the count number of unique calendar year for each subject. All these variables will be used for following checking.

We also merged the enroll information (`firstADRDyear`) to `temp`. (`firstADRDyr`+1) indicates the year that subjects should start to be followed-up.

`temp` is a one-row-per-person dataset, looks like:

```{r echo=FALSE}
head(temp)
```

```{r echo=FALSE}
cat("the number of subjects in ADRDpeople(removed NAs) is", dim(temp)[1], "\n")
cat("the number of person-years in ADRDpeople(removed NAs) is", dim(ADRDpeople)[1], "\n")
cat("is there any duplication of the combination of `qid` and calendar year: ", any(duplicated(ADRDpeople[,.(qid,year)])), "\n")
```

### 2. check whether all people were followed-up from the year following firstADRDyr
```{r}
temp[start_yr != (firstADRDyr+1)]
```

The above subject (`r dim(temp[start_yr != (firstADRDyr+1)])[1]` subjects in total), were not followed-up from the year following firstADRDyr. **Consider deleting them**

Their info in denominator files:
```{r echo=FALSE}
ADRDpeople[qid %in% temp[start_yr != (firstADRDyr+1)][, qid], ]
```

```{r echo=FALSE}
cat("the number of person-years of related subjects is", dim(ADRDpeople[qid %in% temp[start_yr != (firstADRDyr+1)][, qid], ])[1], "\n")
```

### 3. check whether alive people were followed-up till the end of study period (2016)
```{r}
temp[qid %in% ADRDpeople[!(dead),qid], ][, end_yr] %>% table()
```

We could see loads of alive subjects weren't followed-up till 2016. This should be considered as right-censored subjects during the analyses.

### 4. check whether all people have each year's info during follow-up
```{r}
temp[(end_yr-start_yr+1) != count,]
```

The above subject (`r dim(temp[(end_yr-start_yr+1) != count,])[1]` subjects in total) do not have each year's info during follow-up. **Consider deleting them**

Their info in denominator files:
```{r echo=FALSE}
ADRDpeople[qid %in% temp[(end_yr-start_yr+1) != count,][,qid],]
```

